Supervised Dimensionality Reduction with Random Projection-Based Ensemble Learning

### 2021-05-16 paper/main.tex

@@

\section{Simulations}
\subsection{Risk versus discovery: an illustrative example}
It is natural to ask under what circumstances we may recover a classifier that performs very well, but fails to discover the true subspace. To elucidate one example where the relationship between the performance of the base/ensembled classifiers and the discovery rate (power) is subtle, we return to the linear discriminant analysis setting. We restrict to two dimensions, so that $X$ has one 'signal' covariate and one 'noise' covariate. We may then directly control the signal-to-noise ratio via the parameter $\Delta$. In Figure~\ref{fig:ChangingDelta} we investigate, by simulation, the effect of varying $\Delta$ on the average power of a single choice of projection ($B_1 = 1$), when we use projections with $d=1$ and an LDA base classifier on a training set of 200 samples, with selection determined by empirical risk on the training set.

\begin{figure}[htbp]
    \centering
    \includegraphics[width=4in]{../results/figures/ChangingDelta.png}
    \caption{The relationship between signal-to-noise ratio and discovery rate.\label{fig:ChangingDelta}}
\end{figure}

We see that, as expected, too low a signal-to-noise ratio $\Delta$ leads to low-powered projections. We also, perhaps less immediately obviously, see that too high a signal-to-noise ratio results in a weakened ability to discriminate good from bad projections (as all projections begin to offer effective classifiers). The important takeaway is that it is not the performance of base classifiers that determines high-powered discovery, but rather how strongly the performance of those classifiers depends on selecting good projections.

For a (very) simple case, we can begin to pick apart what's going on. Suppose, as above, we have 2-dimensional $X$ following an LDA model. Suppose we have access to an infinite training set (i.e for each projection $A$ we form the Bayes classifier $C_A^{\text{Bayes}}$), but only one validation example. Furthermore, suppose we restrict to $B_2$ = 2, but also impoose that our $d=1$ dimensional projections are not only normalised, but orthogonal. In particular, we may write $A_1^T = (\cos(\theta), \sin(\theta)), A_2^T = (-\sin(\theta), \cos(\theta))$, where $\theta$ will be chosen uniformly from $[0, 2\pi)$. Given our validation sample $(X, Y)$, we have
\begin{align*}
    \hat{R}_1 & = \mathbbm{1}\{ A_1X \geq 0, Y = 0 \} + \mathbbm{1}\{ A_1X \leq 0, Y = 1 \} \\
    \hat{R}_2 & = \mathbbm{1}\{ A_2X \geq 0, Y = 0 \} + \mathbbm{1}\{ A_2X \leq 0, Y = 1 \}
\end{align*}
Since $A_1X$ is independent of $A_2X$ (due to the orthogonality of $A_1, A_2$), we can compute 
\begin{align*}
    \mathbb{P}(\hat{R}_1 \geq \hat{R}_2) & = (1 - \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))\Phi\big(\frac{\Delta}{2}|\sin\theta|\big) \\
    \mathbb{P}(\hat{R}_1 \leq \hat{R}_2) & = \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))(1-\Phi\big(\frac{\Delta}{2}|\sin\theta|\big)) %\\
    % \mathbb{P}(\hat{R}_1 = \hat{R}_2) & = (1 - \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))(1-\Phi\big(\frac{\Delta}{2}|\sin\theta|\big)) + 
    % \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))\Phi\big(\frac{\Delta}{2}|\sin\theta|\big)
\end{align*}
from which we compute the expected power of the selected projection $A_* = A_{\argmin \{ \hat{R}_i\}}$ as 
\begin{align*}
\mathbb{E}[PW(A_*)] = & \mathbb{E}_{\theta \sim \text{Unif}[0, 2\pi)}[\cos^2(\theta) (1 - \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))\Phi\big(\frac{\Delta}{2}|\sin\theta|\big)\\ 
 &+ \sin^2(\theta) \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))(1-\Phi\big(\frac{\Delta}{2}|\sin\theta|\big) \\
 &+ \frac{1}{2} \Big( (1 - \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))(1-\Phi\big(\frac{\Delta}{2}|\sin\theta|\big)) + 
    \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))\Phi\big(\frac{\Delta}{2}|\sin\theta|\big) \Big)].  
    \end{align*}
We plot this curve in Figure~\ref{fig:} -- note the similarity to Figure~\ref{fig:ChangingDelta}.
\subsection{Known Subspace Dimension}

>>>>

\subsection{General Case}
For a projection $A$, let our risk estimate be given by $\hat{R}_A$, where $\hat{R}_A$ takes values in a set $\mathcal{R}$ (for example we may have for validation error $\mathcal{R} = \{0, \frac{1}{n_{\text{val}}}, \frac{2}{n_{\text{val}}}, \ldots, 1 \} $). We assume that $\mathcal{R}$ is finite\footnote{\jbnote{I don't think finite's necessary but pah}} and has at least two distinct values $r_1, r_2 \in \mathcal{R}$ such that $\mathbb{P}(\hat{R}_A = r_i) \neq 0$. We then define the function\footnote{\jbnote{This is expectation over random A, but I don't know how to bold blackboard bold.}} 
\[ \rho(r) := \mathbb{E}[PW(A) | \hat{R}_A = r]. \]
If $\rho$ is strictly decreasing, then for $B2 > 1$:
\[ \mathbb{E}[PW(A_{b_1^*})] > \mathbb{E}[PW(A_{b_1b_2})]. \]
This is since
\begin{align*}

\end{align*}

@@

\section{Simulations}
\subsection{Risk versus discovery: an illustrative example}
It is natural to ask under what circumstances we may recover a classifier that performs very well, but fails to discover the true subspace. To elucidate one example where the relationship between the performance of the base/ensembled classifiers and the discovery rate (power) is subtle, we return to the linear discriminant analysis setting. We restrict to two dimensions, so that $X$ has one 'signal' covariate and one 'noise' covariate. We may then directly control the signal-to-noise ratio via the parameter $\Delta$. In Figure~\ref{fig:ChangingDelta}A we investigate, by simulation, the effect of varying $\Delta$ on the average power of a single choice of projection ($B_1 = 1$), when we use projections with $d=1$ and an LDA base classifier on a training set of 200 samples, with selection determined by empirical risk on the training set.

\begin{figure}[htbp]
    \centering
    \includegraphics[width=5in]{../results/figures/ChangingDelta.png}
    \caption{The relationship between signal-to-noise ratio and discovery rate.\label{fig:ChangingDelta}}
\end{figure}

We see that, as expected, too low a signal-to-noise ratio $\Delta$ leads to low-powered projections. Perhaps less immediately obvious, we also see that too high a signal-to-noise ratio results in a weakened ability to discriminate good from bad projections (as all projections begin to offer effective classifiers).

We can begin to pick apart what's going on by focusing in on an even simpler case. Suppose that, as above, we have 2-dimensional $X$ following an LDA model, but now we have access to an infinite training set (i.e for each $A$ we form the projected Bayes classifier $C_A^{\text{Bayes}}$), but only one validation example. Furthermore, suppose we restrict to $B_2$ = 2, but also impose that our 1-dimensional projections are not only normalised, but orthogonal. In particular, we may write $A_1^T = (\cos(\theta), \sin(\theta)), A_2^T = (-\sin(\theta), \cos(\theta))$, where $\theta$ will be chosen uniformly from $[0, 2\pi)$. Given our validation sample $(X, Y)$, we have
\begin{align*}
    \hat{R}_1 & = \mathbbm{1}\{ A_1X \geq 0, Y = 0 \} + \mathbbm{1}\{ A_1X \leq 0, Y = 1 \} \\
    \hat{R}_2 & = \mathbbm{1}\{ A_2X \geq 0, Y = 0 \} + \mathbbm{1}\{ A_2X \leq 0, Y = 1 \}
\end{align*}
Since $A_1X$ is independent of $A_2X$ (due to the orthogonality of $A_1, A_2$), we can compute 
\begin{align*}
    \mathbb{P}(\hat{R}_1 \geq \hat{R}_2) & = (1 - \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))\Phi\big(\frac{\Delta}{2}|\sin\theta|\big) \\
    \mathbb{P}(\hat{R}_1 \leq \hat{R}_2) & = \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))(1-\Phi\big(\frac{\Delta}{2}|\sin\theta|\big)) %\\
    % \mathbb{P}(\hat{R}_1 = \hat{R}_2) & = (1 - \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))(1-\Phi\big(\frac{\Delta}{2}|\sin\theta|\big)) + 
    % \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))\Phi\big(\frac{\Delta}{2}|\sin\theta|\big)
\end{align*}
from which we compute the expected power of the selected projection $A_* = A_{\argmin \{ \hat{R}_i\}}$ as 
\begin{align*}
\mathbb{E}[PW(A_*)] = & \mathbb{E}_{\theta \sim \text{Unif}[0, 2\pi)}[\cos^2(\theta) (1 - \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))\Phi\big(\frac{\Delta}{2}|\sin\theta|\big)\\ 
 &+ \sin^2(\theta) \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))(1-\Phi\big(\frac{\Delta}{2}|\sin\theta|\big) \\
 &+ \frac{1}{2} \Big( (1 - \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))(1-\Phi\big(\frac{\Delta}{2}|\sin\theta|\big)) + 
    \Phi\big(\frac{\Delta}{2}|\cos\theta|\big))\Phi\big(\frac{\Delta}{2}|\sin\theta|\big) \Big)].  
    \end{align*}
We plot this curve in Figure~\ref{fig:ChangingDelta}B -- note the similarity to \ref{fig:ChangingDelta}A. An important takeaway is that it is not the performance of base classifiers that determines high-powered discovery, but rather how strongly the performance of those classifiers depends on selecting good projections.
\subsection{Known Subspace Dimension}
